\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{physics}
\usepackage{siunitx}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{hyperref}

\title{The Solar System HW3}
\author{Mediumunwell}
\date{\today}

\begin{document}
\maketitle

\section*{Problem 2}
For this question, you will use the Solar System Collisions website at
\url{http://janus.astro.umd.edu/astro/impact/}.
\begin{enumerate}[label=(\alph*)]
\item Investigate collisions with the planets Mars, Venus, and Earth. Determine the maximum-sized
rocky object that is destroyed in the planet's atmosphere to two significant figures.
Use the default collision speed of $\SI{20}{km/s}$.
\item How much energy is released by the largest airbursts on each planet (in Megatons)?
How often does this happen?
\item What are the smallest craters that can be produced on these planets?
\item Saturn's satellite Titan has an atmosphere about ten times thicker than Earth's.
What sorts of impact craters might you expect to find on its surface?
\end{enumerate}

\subsection*{Solution}
I used the simulator with the default speed of $\SI{20}{km/s}$ and recorded the transition values.

\begin{center}
\begin{tabular}{p{0.34\textwidth}p{0.2\textwidth}p{0.2\textwidth}p{0.2\textwidth}}
\hline
Quantity & Mars & Venus & Earth \\
\hline
Largest fully destroyed rocky object & 1.6 m & 6.5 km & 64 m \\
Largest-airburst energy (Mt) & $2.44\times10^{-4}$ & $1.6\times10^{7}$ & 16 \\
Event frequency & every 4.2 months & every 23 million years & every 350 years \\
Smallest crater size & 28 m dia, 5.5 m deep & 83.4 km dia, 1.2 km deep & 836 m dia, 163 m deep \\
\hline
\end{tabular}
\end{center}

For Mars, a 1.6 m object is still fully destroyed in the atmosphere. At 1.7 m, the simulator starts
reporting a crater (28 m diameter, 5.5 m depth), so I used that as the crater threshold.

For Venus, 6.5 km is still fully destroyed, while 6.6 km produces a crater (83.4 km diameter,
1.2 km depth), so that is the first crater-forming case.

For Earth, my earlier value was wrong because I was effectively checking comet-like behavior.
Using the rocky-object setup, the simulator gives 64 m as the largest fully destroyed case,
with 16 Mt and an event frequency of about every 350 years.

For Titan (part d), I would expect many fewer small craters than on airless bodies, because the
thicker atmosphere should destroy or strongly slow small impactors before they reach the ground.
The craters that do form should be biased toward larger impactors.

\section*{Problem 3}
Pressure at Earth's center.
\begin{enumerate}[label=(\alph*)]
\item Find the accepted value from Appendix E. Compare with the predictions of a uniform
density planet by first deriving and then applying Eq. 3.21. Finally, work out the fraction
of the pressure that comes from the core, the part of the planet with $r < R/2$.
\item Make a more realistic approximation by considering a two-component Earth with an iron core
mass $m_c = 1.98 \times 10^{24}\,\si{kg}$ and radius $r_c = \SI{3485}{km}$ and a mantle
mass $m_m = 4.00 \times 10^{24}\,\si{kg}$ and radius $r_m = \SI{6371}{km}$. Determine average
core and mantle densities. Use linearly increasing gravity in the core and approximately constant
gravity in the mantle. Find the gravity at the core-mantle boundary, compare with surface gravity,
then compute and sum the core and mantle contributions to central pressure.
\end{enumerate}

\subsection*{Solution}
\textbf{(a) Uniform-density Earth and comparison to Appendix E}

Start from hydrostatic equilibrium:
\[
\frac{dP}{dr}=-\rho g(r)
\]
and for uniform density,
\[
M(r)=\frac{4\pi}{3}\rho r^3,
\qquad
g(r)=\frac{GM(r)}{r^2}=\frac{4\pi G\rho}{3}r
\]
so
\[
\frac{dP}{dr}=-\frac{4\pi G\rho^2}{3}r
\]
Integrate from radius $r$ to the surface $R$ with $P(R)=0$:
\[
P(r)=\int_r^R \rho g(r')\,dr'
=\frac{2\pi G\rho^2}{3}(R^2-r^2)
\]
Therefore at the center:
\[
P_c=\frac{2\pi G\rho^2R^2}{3}
=\frac{3GM^2}{8\pi R^4}
\]

Using $\rho\approx\SI{5514}{kg/m^3}$ and $R=\SI{6.371e6}{m}$:
\[
\boxed{P_{c,\mathrm{uniform}}\approx\SI{1.73e11}{Pa}=\SI{173}{GPa}}
\]

Accepted value (Appendix E):
\[
\boxed{P_{c,\mathrm{accepted}}\approx\SI{3.6e11}{Pa}=\SI{360}{GPa}}
\]

Comparison:
\[
\frac{P_{c,\mathrm{accepted}}}{P_{c,\mathrm{uniform}}}\approx\frac{360}{173}\approx2.1
\]
so the uniform model underestimates the true central pressure by about a factor of 2.

Pressure fraction from $r<R/2$:
\[
\frac{\int_0^{R/2}\rho g\,dr}{\int_0^R\rho g\,dr}
=\frac{\int_0^{R/2}r\,dr}{\int_0^R r\,dr}
=\frac{(R/2)^2}{R^2}=\frac14
\]
\[
\boxed{\text{Inner half contributes }25\%\text{ of }P_c}
\]

Limit/consistency checks:
\[
P(R)=\frac{2\pi G\rho^2}{3}(R^2-R^2)=0
\]
and $P_c\propto \rho^2R^2$, so denser/larger planets have larger central pressure, which is physically sensible.

\vspace{\baselineskip}


\textbf{(b) Two-component Earth model}

Given:
\[
m_c=1.98\times10^{24}\,\si{kg},\quad r_c=\SI{3485}{km},\quad
m_m=4.00\times10^{24}\,\si{kg},\quad r_m=\SI{6371}{km}
\]

Average densities:
\[
\rho_c=\frac{m_c}{\frac43\pi r_c^3}\approx\SI{1.12e4}{kg/m^3},
\qquad
\rho_m=\frac{m_m}{\frac43\pi(r_m^3-r_c^3)}\approx\SI{4.42e3}{kg/m^3}
\]

Core-mantle boundary and surface gravity:
\[
g_{\mathrm{cmb}}=\frac{Gm_c}{r_c^2}\approx\SI{10.88}{m/s^2},
\qquad
g_s=\frac{G(m_c+m_m)}{r_m^2}\approx\SI{9.83}{m/s^2}
\]
These are close, so using an average mantle gravity is reasonable:
\[
g_{\mathrm{avg}}=\frac{g_{\mathrm{cmb}}+g_s}{2}\approx\SI{10.36}{m/s^2}
\]

Core contribution (with linear $g$ in core):
\[
\Delta P_{\mathrm{core}}
=\int_0^{r_c}\rho_c\left(\frac{Gm_c}{r_c^3}r\right)dr
=\frac{\rho_cGm_c}{2r_c}
\approx\SI{2.12e11}{Pa}=\SI{212}{GPa}
\]

Mantle contribution (constant-$g$ approximation):
\[
\Delta P_{\mathrm{mantle}}
\approx\int_{r_c}^{r_m}\rho_m g_{\mathrm{avg}}\,dr
=\rho_m g_{\mathrm{avg}}(r_m-r_c)
\approx\SI{1.32e11}{Pa}=\SI{132}{GPa}
\]

Total central pressure:
\[
P_{c,\mathrm{two\mbox{-}layer}}
\approx\Delta P_{\mathrm{core}}+\Delta P_{\mathrm{mantle}}
\approx\SI{3.44e11}{Pa}=\SI{344}{GPa}
\]
\[
\boxed{P_{c,\mathrm{two\mbox{-}layer}}\approx\SI{344}{GPa}}
\]

This is much closer to the accepted $\sim\SI{360}{GPa}$ value.

Limit/physical checks:
if $m_c\to0$, then $\Delta P_{\mathrm{core}}\to0$ and the model reduces to a mantle-only case.
If $r_c\to0$, the core region vanishes, again giving no core contribution.
Both limits behave the way they should physically.

\newpage
\section*{Problem 4}
Impact pressure.
\begin{enumerate}[label=(\alph*)]
\item The pressure produced by an impact depends on speed $v$, impactor radius $R$, and density
$\rho$. Use dimensional analysis for
\[
P=C\rho^xR^yv^z
\]
\item Determine $C$ using momentum flux arguments for a normally incident impact.
\item Evaluate $P$ for $R=\SI{10}{km}$, $\rho=\SI{3}{g/cm^3}$,
$v=\SI{20}{km/s}$ and compare with Earth's central pressure.
\end{enumerate}

\subsection*{Solution}
\textbf{(a) Dimensional analysis}

Dimensions:
\[
[P]=M L^{-1}T^{-2},\quad [\rho]=ML^{-3},\quad [R]=L,\quad [v]=LT^{-1}
\]
So
\[
ML^{-1}T^{-2}=(ML^{-3})^xL^y(LT^{-1})^z
=M^xL^{-3x+y+z}T^{-z}
\]
Match exponents:
\[
x=1,\quad z=2,\quad -3x+y+z=-1\Rightarrow y=0
\]
Therefore
\[
\boxed{x=1,\ y=0,\ z=2},\qquad P=C\rho v^2
\]

\textbf{(b) Solve for $C$ from momentum flux}

For $d\ll R$, in time $dt$ the impactor sweeps target mass
\[
dm=\rho A\,v\,dt
\]
with frontal area $A=\pi r^2$.
Accelerating that mass from $0$ to $v$ gives
\[
d p = v\,dm \Rightarrow
F=\frac{dp}{dt}=\rho A v^2
\]
Therefore
\[
P=\frac{F}{A}=\rho v^2
\]
so
\[
\boxed{C=1}
\]

Limit check:
if $v\to0$ or $\rho\to0$, then $P\to0$, which matches physical intuition.

\textbf{(c) Numerical evaluation and comparison}

Numerical value:
\[
\rho=\SI{3}{g/cm^3}=\SI{3000}{kg/m^3},\quad v=\SI{2.0e4}{m/s}
\]
\[
P=\rho v^2=3000(2.0\times10^4)^2
=\SI{1.2e12}{Pa}=\SI{1200}{GPa}
\]
\[
\boxed{P_{\mathrm{impact}}\approx\SI{1200}{GPa}}
\]
Compared with Earth's center pressure $\sim\SI{360}{GPa}$:
\[
\frac{P_{\mathrm{impact}}}{P_{c,\oplus}}\approx\frac{1200}{360}\approx3.3
\]
Thus impact-generated pressures can exceed deep-interior pressures, consistent with formation
of high-pressure mineral phases and shock metamorphism near large impact sites.

\end{document}